---
title: "Week 3 Assignment"
author: "Warner Alexis"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


Please deliver links to an R Markdown file (in GitHub and rpubs.com) with solutions to the problems below.  You may work in a small group, but please submit separately with names of all group participants in your submission.


#1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"



```{r, echo=FALSE}
library(tidyverse)
library(stringr)
library(gt)
```

I am going to provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r, major}
# read the data in GithHub
major <-  read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv",
                  stringsAsFactors = F, header = T)
data_m <- major %>% filter(str_detect(Major, "DATA|STATISTICS"))

gt_data <- gt(data_m)
# Create two additional footnotes, using the
# `columns` and `where` arguments of `data_cells()`
gt_data |>
  tab_header(
    title = "The Data Science and Technology Majors",
    subtitle = "The Only DATA and STATITICS Majors"
  )

# Show the gt Table
gt_data



```

## 2 Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

nto a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")


```{r}
str <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

str
```

## 3 Describe, in words, what these expressions will match:
(.)\1\1
"(.)(.)\\2\\1"
(..)\1
"(.).\\1.\\1"
"(.)(.)(.).*\\3\\2\\1"


```{r}
abc <- c("abc\1", "a\1", "abc","z\001\001","Z\1\1","b\1\1","aaaa","aabbbbbcccc","dd")
v <- "(.)\1\1"
v
str_detect(abc,v)
str_match(abc,v)
```
The backreference \1 (backslash one) references the first capturing group. \1 matches the exact same text that was matched by the first capturing group. The / before it is a literal character. 

"(.)\1\1" will only match group strings group characters follow by \1 

According to the R for Data Science book, The first way to use a capturing group is to refer back to it within a match with back reference: \1 refers to the match contained in the first parenthesis, \2 in the second parenthesis, and so on


```{r}
d <- "(.)(.)\\2\\1"
str_detect(abc,d)
str_match(abc,d)
```
"(.)(.)\\2\\1" would match any match contain in the second parenthesis which mean it will match any four of the same letters. Examples: "aaaa","aabbbbbcccc"--- It will only match a and b 

```{r}
j <- c("(..)\1")
str_detect(abc,j)
str_match(abc,j)
```

"(..)\1" will match any of the last two characters in a group string folowed by "\1". For Example: 
"abc\1" -- bc will be selected 
"acgdeftstwrhyg9.\1" --- g9 will be selected.

```{r}
c_l <- c("cdcacdabbb11","dgdgdfg","abacgwabda","trtrtrtr")
p <- c("(.).\\1.\\1")
str_detect(c_l,p)
str_match(c_l,p)
```


"(.).\\1.\\1" will match string characters only where their first letter is identical after every other string characters. For example: 
"cdcacdabbb11" will match "c" 
"trtrtrtr" will match "t"


```{r}
ch <- c("bcdbcbdbcbd","cdcacdabbb11","dfdfhjdfh","spsdlkjspsd","000550050005")
l <- c("(.)(.)(.).*\\3\\2\\1")
str_detect(ch,l)
str_match(ch,l)
```

"(.)(.)(.).*\\3\\2\\1" will only match characters that repeat three times in a string group. 
For Examples : 
 "dbcbdbcbd" will macth dbc
 "spsdlkjsps"  will match "s"  "p"  "s" 
 
 
 ## 4 Construct regular expressions to match words that:
Start and end with the same character.
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)


```{r}
words <-  c("alababa","cardiac", "chaotic","clementine","blueberry","guava","jujube" )
str_view(words,  "^(.).*\\1$",match = T)
```

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

```{r}
v <- c("(.).*\\1.*\\1")
str_detect(words,v)
str_view(words, v, match = T)

```
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)


```{r}
k <- ("(.).*\\1.*\\1")
str_match(fruit,k)
```

